Defining Transcriptome using RNA Sequencing\\ of {\em Streptococcus mutans}

### 2012-04-05 package/manuscript.Rnw
Just an update.

<<echo=FALSE,result=hide>>=
ua159only <- newSmutans( smutansGenes, title="UA159 Only" )
ua159only <- smutans.de2( ua159only, type="ua159", 
                          condA="glucose", condB="galactose" )
save(ua159only, file=file.path("ua159only.RData"))
#load("ua159only.RData")
x <- smutans.de2List( ua159only, qval=1e-10 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","Glucose","Galactose","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableUA159Only <- x
@
Among the four pairwise comparisons this showed the largest number of
differentially expressed genes (table \ref{tabUA159Only}). We used a smaller
cutoff value for listing them. 
Figure \ref{figPieUA159Only} shows proportons of functional categories that
included differentially expressed genes between glucose-grown and
galactose-grown UA159 wild-type cells. Qualitatively, energy metabolism and PTS
classes were profound. 
``Energy metabolism; Fermentation'' had highly differentially expressed genes
(fig.\ \ref{figDotplotUA159Only})
This was silmilar to the comparison between UA159
wild-type and TW1 mutant-type grown in glucose. 
More than a quarter of the
differentially expressed genes were unknown, unassigned, or hypothetical.
Consecutive genes (SMU.1488c to SMU.1496) were upregulated in 
catabolite-derepressing conditions or in galactose-grown cells
(\fixme{Are these the genes of the lactose (lac) operon?})
(table \ref{tabUA159Only}).  Another set of genes (SMU.886 to SMU.889) were also
upregulated (\fixme{Are these called galKTE?}). Whereas the
four genes appeared to be transcribed together under galactose condition,
two segments seemed to be transcribed by dividing the four genes into two pairs
of genes (fig.\ \ref{figUCSC2}).  Genes from SMU.1535 to SMU.1539, which
appeared to be transcribed together, were upregulated.
Of the \Sexpr{nrow(tableUA159Only)} differentially expressed genes at
$1 \times 10^{-10}$ cutoff of multiple testing adjusted p-value, 
\Sexpr{length(grep("hypothetical",tableUA159Only[,"Gene Description"]))} were
hypothetical, and most of them were down-regulated
(fig.\ \ref{figDotplotUA159Only}).
We also tested functional category enrichments using gene ontology terms 
(table \ref{tab:goseqUA159Only}) and KEGG pathways 
(table \ref{tab:kegseqUA159Only})
where carbohydrate metabolic process and Galactose metabolism were top listed,
respectively.

>>>>

<<echo=FALSE,result=hide>>=
ua159only <- newSmutans( smutansGenes, title="UA159 Only" )
ua159only <- smutans.de2( ua159only, type="ua159", 
                          condA="glucose", condB="galactose" )
save(ua159only, file=file.path("ua159only.RData"))
#load("ua159only.RData")
x <- smutans.de2List( ua159only, qval=1e-10 )
x <- x[,c("id","baseMeanA","baseMeanB","log2FoldChange","padj")]
colnames(x) <- c("Gene ID","Glucose","Galactose","fold", "p-value")
x <- merge(x, geneDescription, by.x = "Gene ID", by.y = "Gene ID")
x <- merge(x, functionalDescription, by.x = "Gene ID", by.y = "Gene ID")
tableUA159Only <- x
@
Among the four pairwise comparisons this showed the largest number of
differentially expressed genes (table \ref{tabUA159Only}). We used a smaller
cutoff value for listing them. 
Figure \ref{figPieUA159Only} shows proportons of functional categories that
included differentially expressed genes between glucose-grown and
galactose-grown UA159 wild-type cells. Qualitatively, energy metabolism and PTS
classes were profound. 
``Energy metabolism; Fermentation'' had highly differentially expressed genes
(fig.\ \ref{figDotplotUA159Only}).
This was similar to the comparison between UA159
wild-type and TW1 mutant-type grown in glucose. 
More than a quarter of the
differentially expressed genes were unknown, unassigned, or hypothetical.
Consecutive genes (SMU.1488c to SMU.1496) were upregulated in 
catabolite-derepressing conditions or in galactose-grown cells
(\fixme{Are these the genes of the lactose (lac) operon?})
(table \ref{tabUA159Only}).  Another set of genes (SMU.886 to SMU.889) were also
upregulated (\fixme{Are these called galKTE?}). Whereas the
four genes appeared to be transcribed together under galactose condition,
two segments seemed to be transcribed by dividing the four genes into two pairs
of genes (fig.\ \ref{figUCSC2}).  Genes from SMU.1535 to SMU.1539, which
appeared to be transcribed together, were upregulated.
Of the \Sexpr{nrow(tableUA159Only)} differentially expressed genes at
$1 \times 10^{-10}$ cutoff of multiple testing adjusted p-value, 
\Sexpr{length(grep("hypothetical",tableUA159Only[,"Gene Description"]))} were
hypothetical, and most of them were down-regulated
(fig.\ \ref{figDotplotUA159Only}).
We also tested functional category enrichments using gene ontology terms 
(table \ref{tab:goseqUA159Only}) and KEGG pathways 
(table \ref{tab:kegseqUA159Only})
where carbohydrate metabolic process and Galactose metabolism were top listed,
respectively.
